---
title: "Project Two"
output:
  pdf_document: default
  html_document: default
---

Due Oct. 21 at 11:59 PM. 

For this first part of the exam, you can either use `surveys_complete.csv` or your own data. If you are using your own data, you must have data in which you think you have a numerical predictor variable and a numerical response variable. If you are using `surveys_complete`, you can use weight and hindfoot_length for this.

Save this file in your `projects` directory. You can either save it in a project two subdirectory, or just put it in projects. Either way is fine.


1) Load in your data. Which variable will you be using as a predictor, and which as a response? (5 pts)

```{r}
# read in data here
surveys <- read_csv("data/portal_data_joined.csv")
surveys_complete <- surveys %>%
     filter(!is.na(weight) +          
            !is.na(hindfoot_length)+  
            !is.na(sex)) 
surveys_complete <- surveys_complete %>%
+     filter(species_id %in% species_counts$species_id)
write_csv(surveys_complete, path = "projects/surveys_complete.csv")
```

```
# Answer which column is predictor and which is response

predictor = weight,
response = hindfoot_length
```

2) Plot the two against each other with a scatter plot. Do the data appear to be related linearly? (5 pts)


```{r}
# Plot here
ggplot(surveys_complete, aes(x = weight, y = hindfoot_length))+ 
+ geom_point()
```

```
#Answer here = NO(none linear)
```


3) Fit the linear model. View the summary. (5 pts)


```{r}
# Code here
ggplot(surveys_complete, aes(x = weight, y = hindfoot_length))+ 
           geom_point()+
           geom_smooth(method = "lm", color = "blue", size = 0.5, fill = "orange")+ 
           labs( x= "Weight", y = "Hindfoot_length", title = "Weight vs hindfoot_length Linear model")+
      annotate("text",x = 20, y = 60, label = "R^2 == 0.4676", parse=T, color = "red", size = 5)
```

4) Does the summary make sense when you compare it to the plot you made? Does our model have good predictive power? Evaluate the residual standard error, intercept, and R-Squared in particular. Would you say your predictor predicts the response?  (10 pts)


```
# Answer here
yes, the summary does ake sense when I compare it to the plot.The model does not have good predictive power due to the R^2 comparitive.I would not say the predictor is a good way predicting this model.
```


5) Plot the model on the graph. Increase the size of the text so it is comfortably readable at 5 feet. Make sure axis labels are readable and not overlapping with one another. (5 pts)

```
# Code here
ggplot(surveys_complete, aes(x = weight, y = hindfoot_length))+ 
           geom_point()+
           geom_smooth(method = "lm", color = "blue", size = 0.5, fill = "orange")+ 
           labs( x= "Weight", y = "Hindfoot_length", title = "Weight vs hindfoot_length Linear model")+
      annotate("text",x = 20, y = 60, label = "R^2 == 0.4676", parse=T, color = "red", size = 5)
```


6) Check the normality of the residuals. Do they look OK? Are we violating assumptions? (5 pts)

```{r}

#Code here
model_fit <- lm(weight ~ hindfoot_length, data = surveys_complete)
 broom::tidy(model_fit)

  broom::augment(model_fit) -> augmented_fit
 qqnorm(augmented_fit$.resid)
 qqline(augmented_fit$.resid, col = "purple")
```

Why is normality of residuals important? 

```{r}

#Answer here
I would say that they do not look ok due to haveing such a abnormal Intercept.
We are violating the normality of the assumption of teh Q-Q plot.
The normality is important because it showes that the prediction is normally distributed.
```

7) If you are using `surveys_complete`: Is there interspecific variation in the linear model? Hint: look at our prior work with facet plots. (15 pts) 

If you are *not* using  `surveys_complete`: Do you think there are groupings within your data that may have a separate linear model? Try grouping the data by that variable and redoing the lm. If this would not make sense for your data, you may also attempt to do multiple predictor variables. (15 pts)

```{r}
#code here
The biggest variation that I see in the two plots is the majority of the points in the facet plot condenst in the middle and then split leading to a sharp r^2 line where the other has a line that runs most of teh points then the points take a sharp upward angle. 
```

## Part Two

In this portion, you are free to use your own data if you have a categorical predictor variable and a response variable. Otherwise, you may use the columns sex and weight in `surveys_complete`

1) First, plot the data grouped by sex (5 pts)

```{r}

# plot code here
ggplot(surveys_complete, aes(x = sex, y = weight)) + geom_jitter()
```

2) Try an ANOVA of this data (5 pt)

```{r}
# ANOVA code here
aov(model_fit) -> anova_model_fit
 summary(anova_model_fit)
```

3) How about a linear model? What information does this give you that an ANOVA can't? (5 pts)


```{r}
#Code here
ggplot(surveys_complete, aes(x = sex, y = weight, color = sex)) + 
     geom_point() + 
     labs(x = "sex",
          y = "weight",
          color = "sex") + 
     scale_color_manual(values = c("blue", "red")) + 
     geom_smooth(method = "lm") +
     annotate("text", x = 20, y = 30, label = "R^2 == 0.4676", parse=T, size=5)
```

```
#Answer here
The linear model showes how the males have a higher number than the females
```

3) Plot the lm with the data, with points colored by sex. (10 pts)


```{r}
#Plot code here
ggplot(surveys_complete, aes(x = sex, y = weight, color = sex)) + 
+     geom_point() + 
+     labs(x = "sex",
+          y = "weight",
+          color = "sex") + 
+     scale_color_manual(values = c("blue", "red")) + 
+     geom_smooth(method = "lm") +
+     annotate("text", x = 20, y = 30, label = "R^2 == 0.4676", parse=T, size=5)
```

4) Choose any model we've looked at so far, but use two predictor variables. Perform an LM, plot the results, and state if this changes your interpretation of the relationship between variables (10 pts)

```{r}
# LM Code Here
model_fit <- lm(sex ~ weight * hindfoot_length, data = surveys_complete)+ summary(model_fit)
```

```{r}
# Plot Code Here 
ggplot(surveys_complete, aes(x = sex, y = weight, color = sex)) + 
     geom_point() + 
     labs(x = "sex",
          y = "weight",
          color = "sex") + 
     scale_color_manual(values = c("blue", "red")) + 
     geom_smooth(method = "lm") +
     annotate("text", x = 20, y = 30, label = "R^2 == 0.4676", parse=T, size=5)
```

```
# Text answer here
No this does not change my interpritation any different.
```

## Part Three


1) Add and commit this document (5 pts)

```
#Commands here
git add <project_two.Rmd>
```

2) Push your changes to github (10 pts)

```
#Commands here
git commit

```



# MS students

My expectation is that you'll do this with your own data. If any part of this doesn't make sense with your data, please get in touch ASAP so we can work it out.
